---
title: "Ellenberg_indicators"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true

knit: (function(input_file, encoding) {
  out_dir <- '../docs';
  rmarkdown::render(input_file,
                    encoding=encoding,
                    output_file=file.path(dirname(input_file), out_dir, 'ellenberg_indicators.html'))})
---
  
```{r setup, include=FALSE, message=FALSE}
library(knitr)
library(tidyr)
library(stringr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(hrbrthemes)
library(viridis)
knitr::opts_chunk$set(echo = TRUE)
```


# Analyses of heat requirement of species in a plant community 
- Data import & handling  
- Building the reference  
- Scaling of the evaluation data to an ecological-condition index  

## Data import & handling
We have three data sources for this indicator:

Ellenberg indicator data  
NiN generalized species lists for mountain ecosystems  
ANO data  
- geography  
- locations  
- species 

```{r data import, include = FALSE}

ell.dat <- read.csv("../data/plant indicators/species indicator data/ellen.csv",sep=";", header=T)

load("../data/plant indicators/NiN/Eco_State.RData")

ANOpkt <- read.table(file="../data/plant indicators/ANO/PunktData_all.txt", dec=',', sep = '\t', header=T)
ANOpkt.geo <- read.table(file="../data/plant indicators/ANO/ANO_geo.csv", sep = ',', header=T)
sp <- read.table(file="../data/plant indicators/ANO/ArtsData_all.txt", dec=',', sep = '\t', header=T)
```

ell.dat is a data frame with British Ellenberg species indicators. Literature reference:

Hill, M. O., Mountford, J. O., Roy, D. B., & Bunce, R. G. H. (1999). Ellenberg's Indicator Values for British Plants. Institute of Terrestrial Ecology, Huntingdon, UK.  
```{r}
head(ell.dat)
```

Eco_State.RData is an R-list containing the generalized species lists from NiN. References:

Halvorsen, R., Skarpaas, O., Bryn, A., Bratli, H., Erikstad, L., Simensen, T., & Lieungh, E. (2020). Towards a systematics of ecodiversity: The EcoSyst framework. Global Ecology and Biogeography, 29(11), 1887-1906. doi:10.1111/geb.13164

Halvorsen, R., Bryn, A., & Erikstad, L. (2015). NiNs systemkjerne – teori, prinsipper og inndelingskriterier. – Natur i Norge, Artikkel 1 (versjon 2.0.0)
```{r eval = FALSE}
str(Eco_State)
```

ANOpkt.geo is a data frame with geographical information for each ANO site and point
```{r}
ANOpkt.geo[1:3,]
```

ANOpkt is a data frame with ecological information for each ANO-point (only showing the header here)
```{r}
colnames(ANOpkt)
```

sp is a data frame with plant species information (occurrence and abundance) for each ANO-point
```{r}
sp[1:3,]
```


```{r data handling, include = FALSE}
#### Data handling
### Indicator data
# trimming away sub-species & co, and descriptor info
ell.dat[,'species'] <- word(ell.dat[,'species'], 1,2)
# dealing with 'duplicates'
ell.dat[duplicated(ell.dat[,'species']),"species"]
ell.dup <- ell.dat[duplicated(ell.dat[,'species']),"species"]
ell.dat[ell.dat$species %in% ell.dup,]
ell.dat <- ell.dat[!duplicated(ell.dat[,'species']),]
ell.dat[ell.dat$species %in% ell.dup,]
ell.dat[duplicated(ell.dat[,'species']),"species"]

## NiN data
# transposing abundance data for bootstrapping
all.dat <- t(Eco_State$Concept_Data$Species$Species_Data)
all.dat <- as.data.frame(all.dat)
all.dat$sp <- as.factor(as.vector(Eco_State$Concept_Data$Species$Species_List$species))
# only genus and species name
all.dat$sp <- word(all.dat$sp, 1,2)
all.dat$spgr <- as.factor(as.vector(Eco_State$Concept_Data$Species$Species_List$art.code))
# environment data
env.dat <- Eco_State$Concept_Data$Env$Env_Data

## merging NiN lists and indicator values
all.ind.dat <- merge(all.dat,ell.dat, by.x="sp", by.y="species", all.x=T)
all.ind.dat[all.ind.dat==999] <- NA

## extracting mountain NiN-lists
mount.dat <- all.ind.dat[,c("sp",paste(env.dat$ID[env.dat$Nature_Type=="Mountain " | env.dat$Sub_Type=="Mountain" | env.dat$Sub_Type=="Mountain"]),"L","N")]
# chosing the relevant species lists and assigning the right NiN names
mount.dat <- mount.dat[,c(1:19,21:26,34:36,40:41)]
colnames(mount.dat)[2:28] <- c('T14-C1','T14-C2','T3-C3C6','T3-C9C12',
                               'T3-C2C5','T3-C8C11','T3-C1a','T3-C1b',
                               'T3-C4','T3-C7','T3-C10','T3-C13',
                               'T3-C14','T7-C1','T7-C2','T7-C3',
                               'T7-C6','T7-C8','T7-C12','T7-C13C14',
                               'T7-C4','T7-C7C9','T7-C5','T7-C10',
                               'T22-C1C3a','T22-C1C3b','T22-C2C4')

# translating the abundance classes into %-cover
coverscale <- data.frame(orig=0:6,
                         cov=c(0,1/32,1/8,3/8,0.6,4/5,1)
)

mount.dat.cov <- mount.dat
for (i in 2:28) {
  mount.dat.cov[,i] <- coverscale[,2][ match(mount.dat[,i], 0:6 ) ]
}


## ANO data

colnames(sp)[5] <- 'Dekning'
# simplify punkt data
pkt.vars <- c("Aar", "GlobalID", "ANO_flate_ID", "ANO_punkt_ID", 
              "Hovedoekosystem_rute", "Hovedoekosystem_sirkel", "Andel_hovedoekosystem_sirkel", 
              "Hovedtype_rute", "Kartleggingsenhet_rute", 
              "Hovedtype_sirkel", "Kartleggingsenhet_sirkel", "Andel_Kartleggingsenhet_sirkel",
              "Arter_tilstede", "Dekning_fremmedarter")

ANOpkt <- ANOpkt[,pkt.vars]
ANOpkt <- ANOpkt[!is.na(ANOpkt$GlobalID),]


# removing duplicate ANOID's (due to multiple historical municipalities)
ANOpkt.geo <- ANOpkt.geo[!duplicated(ANOpkt.geo$ANOID), ]

# fixing names and factor levels in ANOpkt.geo
colnames(ANOpkt.geo)[c(12,16,21)] <- c("Year","Region","Region_no")
ANOpkt.geo$CountyName <- as.factor(ANOpkt.geo$CountyName)
levels(ANOpkt.geo$CountyName) <- c("Oestfold","Akershus",         
                                   "Aust-Agder","Buskerud",         
                                   "Finnmark","Hedmark",          
                                   "Hordaland","Moere og Romsdal",
                                   "Nordland","Oppland",          
                                   "Oslo","Rogaland",         
                                   "Sogn og Fjordane","Telemark",         
                                   "Troendelag","Troms",          
                                   "Vest-Agder","Vestfold") 
ANOpkt.geo$Fylke <- as.factor(ANOpkt.geo$Fylke)
levels(ANOpkt.geo$Fylke) <- c("Oestfold","Akershus",         
                              "Aust-Agder","Buskerud",         
                              "Finnmark","Hedmark",          
                              "Hordaland","Moere og Romsdal",
                              "Nord-Troendelag",
                              "Nordland","Oppland",          
                              "Oslo","Rogaland", 
                              "Hordaland","Moere og Romsdal",
                              "Soer-Troendelag",
                              "Sogn og Fjordane","Telemark",         
                              "Troendelag","Troms",          
                              "Vest-Agder","Vestfold") 
# trimming ANOpkt.geo
colnames(ANOpkt.geo)
ANOpkt.geo <- ANOpkt.geo[,c(4,8,9,12,15,16)]

# merging ANOpkt & ANOpkt.geo
ANOpkt <- merge(x=ANOpkt, y= ANOpkt.geo,by.x="ANO_flate_ID", by.y="ANOID", all.x=T)

## add indicator values to species data
# fix species names
head(sp)
sp$Species <- sp$Navn
sp$Species <- sub(".*?_", "", sp$Species) # lose the Norwegian name in the front
sp$Species <- sub("^([^_]*_[^_]*).*", "\\1", sp$Species) # lose subspecies
sp$Species <- str_to_title(sp$Species) # make first letter capital
sp$Species <- gsub("_", " ", sp$Species) # replace underscore with space
sp$Species <- gsub("(-.*)","\\L\\1",sp$Species,perl=TRUE) # make capital letters after hyphon to lowercase

## merge species data with indicators
sp <- merge(x=sp, y= ell.dat[,c("species","L","N")],by.x="Species", by.y="species", all.x=T)

summary(sp)

## adding hovedøkosystem, hovetype & kartleggingsenhet
sp <- merge(x=sp, y=ANOpkt[,c("GlobalID","Hovedoekosystem_rute","Hovedtype_rute","Kartleggingsenhet_rute")], 
            by.x="ParentGlobalID", by.y="GlobalID", all.x=T)
# trimming away the points without information on NiN, species or cover
sp$Hovedoekosystem_rute <- as.factor(sp$Hovedoekosystem_rute)
sp$Hovedtype_rute <- as.factor(sp$Hovedtype_rute)
sp$Kartleggingsenhet_rute <- as.factor(sp$Kartleggingsenhet_rute)
sp$Species <- as.factor(sp$Species)
sp <- sp[!is.na(sp$Hovedoekosystem_rute),]
sp <- sp[!is.na(sp$Species),]
sp <- sp[!is.na(sp$Dekning),]


# ...which makes us end up with:  
  
# NiN species lists for mountains matched with indicator values  
# Note that the abundance values in the NiN species lists are maximum abundance!
head(mount.dat.cov)
# Species data from ANO matched with indicator values
head(sp)
```
After some merging and data wrangling we arrive at:
(i) NiN species lists for mountains matched with Heat-requirement values (Note that the abundance values in the NiN species lists are maximum abundance). From here we can build the reference.
``` {r}
mount.dat.cov[1:3,]
```

...and at (ii) Species data from the ANO monitoring program matched with Heat-requirement values
``` {r}
sp[1:3,]
```

## Building the reference (scaling values)

### Concept for the reference
The reference condition is defined by a distribution of species indicator values (for mountain ecosystems these are Ellenberg L & N) for each basic ecosystem type (grunntype) in NiN. To get such indicator distributions we bootstrap the generalized species lists underlying the NiN-ecosystem-type-categorization on the grunntype-level and calculate a community weighted mean of Ellenberg L & N for each bootstrap in each generalized species list. That way we generate a reference distribution for each grunntype in NiN that is connected to an underlying generalized species list.  

### Bootstrap function  
``` {r ref bootstrap function}
# function to calculate community weighted means of selected indicator values (ind)
# for species lists (sp) with given abundances (on a scale from 0 to 6) in one or more 'sites' (abun)
# with a given number of iterations (iter),
# with species given a certain minimum abundance occurring in all bootstraps (obl), and
# with a given resampling ratio of the original species list (rat)
# in every bootstrap iteration the abundance of the sampled species can be 
# randomly changed by +/-1 if wished (var.abun)

indBoot <- function(sp,abun,ind,iter,obl,rat=2/3,var.abun=F) {
  
  ind.b <- matrix(nrow=iter,ncol=length(colnames(abun)))
  colnames(ind.b) <- colnames(abun)
  ind.b <- as.data.frame(ind.b)  
  
  ind <- as.data.frame(ind)
  ind.list <- as.list(1:length(colnames(ind)))
  names(ind.list) <- colnames(ind)
  
  for (k in 1:length(colnames(ind)) ) 
    { ind.list[[k]] <- ind.b }
  
  for (j in 1:length(colnames(abun)) ) {
    
    dat <- cbind(sp,abun[,j],ind)
    dat <- dat[dat[,2]>0,]            # only species that are present in the ecoystem
    dat <- dat[!is.na(dat[,3]),]      # only species that have Ellenberg values
    
    for (i in 1:iter) {
      
      speciesSample <- sample(dat$sp[dat[,2] < obl], size=round( (length(dat$sp)-length(dat$sp[dat[,2] > (obl-1)])) *rat,0), replace=F)  
      dat.b <- rbind(dat[dat[,2] > (obl-1),],
                     dat[match(speciesSample,dat$sp),]
                     )
      
      if (var.abun==T) {
        for (l in 1: nrow(dat.b)) { dat.b[l,2] <- dat.b[l,2] + sample(c(-1,0,1),size=1) }
        dat.b[!is.na(dat.b[,2]) & dat.b[,2]==(0),2] <- 1
        dat.b[!is.na(dat.b[,2]) & dat.b[,2]==7,2] <- 6
      }
      
      for (k in 1:length(colnames(ind))) {
        
        if ( nrow(dat.b)>2 ) {
          
          ind.b <- sum(dat.b[!is.na(dat.b[,2+k]),2] * dat.b[!is.na(dat.b[,2+k]),2+k] , na.rm=T) / sum(dat.b[!is.na(dat.b[,2+k]),2],na.rm=T)
          ind.list[[k]][i,j] <- ind.b
          
        } else {ind.list[[k]][i,j] <- NA}
        
      }

            print(paste(i,"",j)) 
    }
    
  }
  return(ind.list)
}

```

### Running the bootstrap function to build the reference 
``` {r running bootstraps, eval = FALSE}
#### running bootstraps ####
mount.ind.list <- indBoot(sp=mount.dat[,1],abun=mount.dat[,2:28],ind=mount.dat[,29:30],
                          iter=10000,obl=6,rat=1/3,var.abun=T)

mount.ind.list.cov <- indBoot(sp=mount.dat.cov[,1],abun=mount.dat.cov[,2:28],ind=mount.dat.cov[,29:30],
                          iter=10000,obl=6,rat=1/3,var.abun=T)


### fixing NaN's
for (i in 1:length(mount.ind.list) ) {
  for (j in 1:ncol(mount.ind.list[[i]]) ) {
    v <- mount.ind.list[[i]][,j]
    v[is.nan(v)] <- NA
    mount.ind.list[[i]][,j] <- v
  }
}

for (i in 1:length(mount.ind.list.cov) ) {
  for (j in 1:ncol(mount.ind.list.cov[[i]]) ) {
    v <- mount.ind.list.cov[[i]][,j]
    v[is.nan(v)] <- NA
    mount.ind.list.cov[[i]][,j] <- v
  }
}

```

``` {r load bootstrapped reference lists, echo = FALSE}
load("../data/plant indicators/NiN/ref_lists_mount211123_10000.RData")
```

### Deriving threshold values
Next we derive the limit and reference values from the bootstrapped distribution.
e.g. upper and lower limit values = 2.5 & 97.5% percentiles, reference value = median, minimum & maximum value given by the minimum and maximum values of the Ellenberg scale (1 & 9).

``` {r derive threshold values, include = FALSE}

# 2 indicator-value indicators: L,N
indEll.n=2
# creating a table to hold:
# Ellenberg: the 0.5 quantile (median), 0.05 quantile and  0.95 quantile for each NiN-type
# for every nature type (nrows)
tab <- matrix(ncol=3*indEll.n, nrow=33+4 ) # 33+4 as there are 33 basic ecosystem types + registrations only stating the four main ecosystem types
# coercing the values into the table
# NiN-types where each type is represented by one species list (including when one species list represents two NiN-types)
names(mount.ind.list.cov[[1]])
x <- c(1:6,9:24,27)

for (i in 1:length(x) ) {
  tab[i,1:3] <- quantile(as.matrix(mount.ind.list.cov[["L"]][,x[i]]),probs=c(0.025,0.5,0.975),na.rm=T)
  tab[i,4:6] <- quantile(as.matrix(mount.ind.list.cov[["N"]][,x[i]]),probs=c(0.025,0.5,0.975),na.rm=T)
}

tab <- as.data.frame(tab)
tab$NiN <- NA
tab$NiN[1:23] <- names(mount.ind.list.cov[[1]])[c(1:6,9:24,27)]
# double rows for species lists representing two NiN-types
mount.ind.list.cov[[1]][0,x]
tab[c(24:30),] <- tab[c(3:6,18,20,23),]
tab$NiN[24:30] <- names(mount.ind.list.cov[[1]])[c(3:6,20,22,27)]
tab
# NiN-types represented by several species lists
tab[31,1:3] <- quantile(as.matrix(mount.ind.list.cov[["L"]][,c(7,8)]),probs=c(0.025,0.5,0.975),na.rm=T)
tab[31,4:6] <- quantile(as.matrix(mount.ind.list.cov[["N"]][,c(7,8)]),probs=c(0.025,0.5,0.975),na.rm=T)
# NiN-types represented by several species lists & double rows for species lists representing two NiN-types
tab[33,1:3] <- tab[32,1:3] <- quantile(as.matrix(mount.ind.list.cov[["L"]][,c(25,26)]),probs=c(0.025,0.5,0.975),na.rm=T)
tab[33,4:6] <- tab[32,4:6] <- quantile(as.matrix(mount.ind.list.cov[["N"]][,c(25,26)]),probs=c(0.025,0.5,0.975),na.rm=T)
tab$NiN[c(31:33)] <- c('T3-C1','T22-C1','T22-C3')
tab
# averages for main NiN-types
# T14
tab[34,1:3] <- quantile(as.matrix(mount.ind.list.cov[["L"]][,c(1,2)]),probs=c(0.025,0.5,0.975),na.rm=T)
tab[34,4:6] <- quantile(as.matrix(mount.ind.list.cov[["N"]][,c(1,2)]),probs=c(0.025,0.5,0.975),na.rm=T)
# T3
tab[35,1:3] <- quantile(as.matrix(mount.ind.list.cov[["L"]][,3:13]),probs=c(0.025,0.5,0.975),na.rm=T)
tab[35,4:6] <- quantile(as.matrix(mount.ind.list.cov[["N"]][,3:13]),probs=c(0.025,0.5,0.975),na.rm=T)
# T7
tab[36,1:3] <- quantile(as.matrix(mount.ind.list.cov[["L"]][,14:24]),probs=c(0.025,0.5,0.975),na.rm=T)
tab[36,4:6] <- quantile(as.matrix(mount.ind.list.cov[["N"]][,14:24]),probs=c(0.025,0.5,0.975),na.rm=T)
# T22
tab[37,1:3] <- quantile(as.matrix(mount.ind.list.cov[["L"]][,25:27]),probs=c(0.025,0.5,0.975),na.rm=T)
tab[37,4:6] <- quantile(as.matrix(mount.ind.list.cov[["N"]][,25:27]),probs=c(0.025,0.5,0.975),na.rm=T)
tab$NiN[c(34:37)] <- c('T14','T3','T7','T22')
tab
# making it a proper data frame
round(tab[,1:6],digits=2)
colnames(mount.ind.list.cov[[1]])
tab$NiN

tab$NiN <- c('T14-C1','T14-C2','T3-C3','T3-C9','T3-C2','T3-C8','T3-C4', 'T3-C7',
             'T3-C10','T3-C13','T3-C14','T7-C1', 'T7-C2', 'T7-C3', 'T7-C6', 'T7-C8', 'T7-C12','T7-C13',
             'T7-C4', 'T7-C7','T7-C5', 'T7-C10',
             'T22-C2',
             'T3-C6','T3-C12','T3-C5','T3-C11','T7-C14','T7-C9',
             'T22-C4',
             'T3-C1',
             'T22-C1','T22-C3',
             'T14','T3','T7','T22')
colnames(tab) <- c("EllLq2.5","EllLq50","EllLq97.5",
                   "EllNq2.5","EllNq50","EllNq97.5",
                   "NiN")
summary(tab)
tab
# restructuring into separate indicators for lower (q2.5) and higher (q97.5) than reference value (=median, q50)
y.L <- numeric(length=nrow(tab)*2)
y.L[((1:dim(tab)[1])*2)-1] <- tab$EllLq2.5
y.L[((1:dim(tab)[1])*2)] <- tab$EllLq97.5

y.N <- numeric(length=nrow(tab)*2)
y.N[((1:dim(tab)[1])*2)-1] <- tab$EllNq2.5
y.N[((1:dim(tab)[1])*2)] <- tab$EllNq97.5

# creating final objects holding the reference and limit values for all indicators

# ref for alien and Ellenberg N
mount.cov.ref <- data.frame(N1= rep('fjell',(nrow(tab)*2*indEll.n)),
                        N2= rep('fjell',(nrow(tab)*2*indEll.n)),
                        hoved= rep('NA',(nrow(tab)*2*indEll.n)),
                        grunn= rep(rep(tab$NiN,each=2),indEll.n),
                        county= rep('all',(nrow(tab)*2*indEll.n)),
                        region= rep('all',(nrow(tab)*2*indEll.n)),
                        Ind= c(rep(c('EllL1','EllL2'),nrow(tab)),rep(c('EllN1','EllN2'),nrow(tab))),
                        Rv= c(rep(tab$EllLq50,each=2),rep(tab$EllNq50,each=2)),
                        Gv=c(y.L,y.N),
                        maxmin=rep(c(1,9),indEll.n*nrow(tab))
                        )

head(mount.cov.ref)



# add extra '-' after 'C' in  mount.cov.ref
mount.cov.ref$grunn <- as.character(mount.cov.ref$grunn)
for (i in 1:nrow(mount.cov.ref) ) {
  if ( grepl('-C',mount.cov.ref$grunn[i]) ) {
    xxx <- as.character(mount.cov.ref$grunn[i])
    n <- which(strsplit(xxx, "")[[1]]=="C")+1
    mount.cov.ref$grunn[i] <- paste(substr(xxx, 1, n-1), "-", substr(xxx, n, nchar(xxx)), sep = "")
  }
}

mount.cov.ref
mount.cov.ref$N1 <- as.factor(mount.cov.ref$N1)
mount.cov.ref$N2 <- as.factor(mount.cov.ref$N2)
mount.cov.ref$hoved <- as.factor(mount.cov.ref$hoved)
mount.cov.ref$grunn <- as.factor(mount.cov.ref$grunn)
mount.cov.ref$county <- as.factor(mount.cov.ref$county)
mount.cov.ref$region <- as.factor(mount.cov.ref$region)
mount.cov.ref$Ind <- as.factor(mount.cov.ref$Ind)
summary(mount.cov.ref)

```

```{r}
summary(mount.cov.ref)
```


## Scaling of the evaluation data to an ecological-condition index
We calculate the relative cumulative cover for species at and above the heat requirement threshold for the ANO data and evaluate these against their respective reference values (i.e. scaling according to reference value=1, limit value=0.6, and maximum value=0)
This is just technical, so we don't show the code.
```{r index scaling, include = FALSE}
#### scaled values ####
r.s <- 1    # reference value
l.s <- 0.6  # limit value
a.s <- 0    # abscence of indicator, or indicator at maximum

#### function for calculating scaled values for measured value ####

## scaling function without truncation
scal.2 <- function() {
  # place to hold the result
  x <- numeric()
  if (maxmin < ref) {
    # values >= the reference value estimated from the linear relationship for lim < x < ref (line below)
    if (val >= ref) {x <- (l.s + (val-lim) * ( (r.s-l.s) / (ref-lim) ) )}
    # values < the reference value and >= the limit value can be deducted from the linear relationship between these two
    if (val < ref & val >= lim) {x <- (l.s + (val-lim) * ( (r.s-l.s) / (ref-lim) ) )}
    # values < the limit value and > maxmin can be deducted from the linear relationship between these two
    if (val < lim & val > maxmin) {x <- (a.s + (val-maxmin) * ( (l.s-a.s) / (lim-maxmin) ) )}
    # value equal or lower than maxmin
    if (val <= maxmin) {x <-0}
  } else {
    # values <= the reference value estimated from the linear relationship for lim < x < ref (line below)
    if (val <= ref) {x <- ( r.s - ( (r.s - l.s) * (val - ref) / (lim - ref) ) )}
    # values > the reference value and <= the limit value can be deducted from the linear relationship between these two
    if (val > ref & val <= lim) {x <- ( r.s - ( (r.s - l.s) * (val - ref) / (lim - ref) ) )}
    # values > the limit value and < maxmin can be deducted from the linear relationship between these two
    if (val > lim & val < maxmin) {x <- ( l.s - (l.s * (val - lim) / (maxmin - lim) ) )}
    # value equal og larger than maxmin
    if (val >= maxmin) {x <-0}
  }
  return(x)
  
}


#### creating dataframe to holde the results for mountains ####
ANOpkt.mount <- ANOpkt[ANOpkt$Hovedoekosystem_rute=='fjell' | ANOpkt$Hovedoekosystem_sirkel=='fjell',]
summary(ANOpkt.mount$Hovedoekosystem_sirkel)
summary(ANOpkt.mount$Hovedoekosystem_rute)
summary(ANOpkt.mount$Hovedtype_rute)
ANOpkt.mount$Hovedtype_rute <- factor(ANOpkt.mount$Hovedtype_rute)
ANOpkt.mount$Kartleggingsenhet_rute <- factor(ANOpkt.mount$Kartleggingsenhet_rute)
summary(ANOpkt.mount$Hovedtype_rute)
summary(ANOpkt.mount$Kartleggingsenhet_rute)


ind <- unique(mount.cov.ref$Ind)
# choose columns for site description
colnames(ANOpkt.mount)
result.ell.mount <- ANOpkt.mount[,c(2,1,3,4:19)]

# add columns for indicators
nvar.site <- ncol(result.ell.mount)
for (i in 1:length(ind) ) {result.ell.mount[,i+nvar.site] <- NA}
colnames(result.ell.mount)[(nvar.site+1):(length(ind)+nvar.site)] <- paste(ind)
for (i in (nvar.site+1):(length(ind)+nvar.site) ) {result.ell.mount[,i] <- as.numeric(result.ell.mount[,i])}
summary(result.ell.mount)
result.ell.mount$Region <- as.factor(result.ell.mount$Region)
result.ell.mount$GlobalID <- as.factor(result.ell.mount$GlobalID)
result.ell.mount$ANO_flate_ID <- as.factor(result.ell.mount$ANO_flate_ID)
result.ell.mount$ANO_punkt_ID <- as.factor(result.ell.mount$ANO_punkt_ID)
result.ell.mount$Hovedoekosystem_sirkel <- as.factor(result.ell.mount$Hovedoekosystem_sirkel)
result.ell.mount$Hovedoekosystem_rute  <- as.factor(result.ell.mount$Hovedoekosystem_rute )
result.ell.mount$Kartleggingsenhet_sirkel <- as.factor(result.ell.mount$Kartleggingsenhet_sirkel)
result.ell.mount$Hovedtype_sirkel    <- as.factor(result.ell.mount$Hovedtype_sirkel)
result.ell.mount$Hovedtype_rute    <- as.factor(result.ell.mount$Hovedtype_rute)


#### calculating scaled and non-truncated values for the indicators based on the dataset ####
for (i in 1:nrow(ANOpkt.mount) ) {  #
  tryCatch({
    print(i)
    print(paste(ANOpkt.mount$ANO_flate_ID[i]))
    print(paste(ANOpkt.mount$ANO_punkt_ID[i]))
#    ANOpkt.mount$Hovedoekosystem_sirkel[i]
#    ANOpkt.mount$Hovedoekosystem_rute[i]



    if (ANOpkt.mount$Hovedtype_rute[i] %in% c('T3','T7','T14','T22') ) {
      
      # Species indicator values  
      if ( length(sp[sp$ParentGlobalID==as.character(ANOpkt.mount$GlobalID[i]),'Species']) > 0 ) {
        
        
          #Ellenberg N
          dat <- sp[sp$ParentGlobalID==as.character(ANOpkt.mount$GlobalID[i]),c('Dekning','N','Kartleggingsenhet_rute')]
          dat <- dat[!is.na(dat$N),]
          
          if ( nrow(dat)>0 ) {
            
            val <- sum(dat[,'Dekning'] * dat[,'N'],na.rm=T) / sum(dat[,'Dekning'],na.rm=T)
            # lower part of distribution
            ref <- mount.cov.ref[mount.cov.ref$Ind=='EllN1' & mount.cov.ref$grunn==as.character(unique(dat$Kartleggingsenhet_rute)),'Rv']
            lim <- mount.cov.ref[mount.cov.ref$Ind=='EllN1' & mount.cov.ref$grunn==as.character(unique(dat$Kartleggingsenhet_rute)),'Gv']
            maxmin <- mount.cov.ref[mount.cov.ref$Ind=='EllN1' & mount.cov.ref$grunn==as.character(unique(dat$Kartleggingsenhet_rute)),'maxmin']
            # coercing x into result.ell.mount dataframe
            result.ell.mount[i,'EllN1'] <- scal.2() 

            
            # upper part of distribution
            ref <- mount.cov.ref[mount.cov.ref$Ind=='EllN2' & mount.cov.ref$grunn==as.character(unique(dat$Kartleggingsenhet_rute)),'Rv']
            lim <- mount.cov.ref[mount.cov.ref$Ind=='EllN2' & mount.cov.ref$grunn==as.character(unique(dat$Kartleggingsenhet_rute)),'Gv']
            maxmin <- mount.cov.ref[mount.cov.ref$Ind=='EllN2' & mount.cov.ref$grunn==as.character(unique(dat$Kartleggingsenhet_rute)),'maxmin']
            # coercing x into result.ell.mount dataframe
            result.ell.mount[i,'EllN2'] <- scal.2() 

          }
          
          #Ellenberg L
          dat <- sp[sp$ParentGlobalID==as.character(ANOpkt.mount$GlobalID[i]),c('Dekning','L','Kartleggingsenhet_rute')]
          dat <- dat[!is.na(dat$L),]
          
          if ( nrow(dat)>0 ) {
            
            val <- sum(dat[,'Dekning'] * dat[,'L'],na.rm=T) / sum(dat[,'Dekning'],na.rm=T)
            # lower part of distribution
            ref <- mount.cov.ref[mount.cov.ref$Ind=='EllL1' & mount.cov.ref$grunn==as.character(unique(dat$Kartleggingsenhet_rute)),'Rv']
            lim <- mount.cov.ref[mount.cov.ref$Ind=='EllL1' & mount.cov.ref$grunn==as.character(unique(dat$Kartleggingsenhet_rute)),'Gv']
            maxmin <- mount.cov.ref[mount.cov.ref$Ind=='EllL1' & mount.cov.ref$grunn==as.character(unique(dat$Kartleggingsenhet_rute)),'maxmin']
            # coercing x into result.ell.mount dataframe
            result.ell.mount[i,'EllL1'] <- scal.2() 
            
            # upper part of distribution
            ref <- mount.cov.ref[mount.cov.ref$Ind=='EllL2' & mount.cov.ref$grunn==as.character(unique(dat$Kartleggingsenhet_rute)),'Rv']
            lim <- mount.cov.ref[mount.cov.ref$Ind=='EllL2' & mount.cov.ref$grunn==as.character(unique(dat$Kartleggingsenhet_rute)),'Gv']
            maxmin <- mount.cov.ref[mount.cov.ref$Ind=='EllL2' & mount.cov.ref$grunn==as.character(unique(dat$Kartleggingsenhet_rute)),'maxmin']
            # coercing x into result.ell.mount dataframe
            result.ell.mount[i,'EllL2'] <- scal.2() 
          }
        }
      }
    
  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}

summary(result.ell.mount)
# remove values >1 for Ellenberg
result.ell.mount$EllL1[result.ell.mount$EllL1>1] <- NA
result.ell.mount$EllL2[result.ell.mount$EllL2>1] <- NA
result.ell.mount$EllN1[result.ell.mount$EllN1>1] <- NA
result.ell.mount$EllN2[result.ell.mount$EllN2>1] <- NA

# only the relevant NiN-types
result.ell.mount <- result.ell.mount[result.ell.mount$Hovedtype_rute %in% c('T3','T7','T14','T22'),]

summary(result.ell.mount)

```
Here's the result
```{r, echo = FALSE}
summary(result.ell.mount[,c('EllL1','EllL2','EllN1','EllN2')])
```

And here's the result as violin plots by main ecosystem types  
T3 = mountain heat, leeside, & tundra (fjellhei, leeside & tundra)  
T14 = ridge (rabbe)  
T7 = snowbed (snøleie)  
T22 = grassy mountain heath & tundra (fjellgrashei & grastundra)  

``` {r result, echo = FALSE}



result.ell.mount2 <- gather(result.ell.mount,value=index,key=indicator,-c(1:19))
result.ell.mount2 <- result.ell.mount2[!is.na(result.ell.mount2$index),]
result.ell.mount2$indicator <- as.factor(result.ell.mount2$indicator)
# sample size
sample_size = result.ell.mount2 %>% group_by(indicator) %>% summarize(num=n())
# add sample size to labels
new_labels <- c("EllL1" = paste(levels(result.ell.mount2$indicator)[1],', N=',sample_size$num[1],sep=""),
                "EllL2" = paste(levels(result.ell.mount2$indicator)[2],', N=',sample_size$num[2],sep=""),
                "EllN1" = paste(levels(result.ell.mount2$indicator)[3],', N=',sample_size$num[3],sep=""),
                "EllN2" = paste(levels(result.ell.mount2$indicator)[4],', N=',sample_size$num[4],sep=""))
# plot
ggplot( result.ell.mount2 , aes(x=Hovedtype_rute, y=index, fill=as.factor(Hovedtype_rute)  )) + 
  geom_violin(width=1.4) +  
  facet_wrap(~indicator , dir="v", labeller = labeller(indicator = new_labels))  +
  theme(legend.position="none") +
    ggtitle("Scaled index results Ellenberg indicators") +
    xlab("Ecosystem type") +
  ylab("Ecological condition index")
```

Ellenberg L shows most deviations from the reference value (index = 1) towards the higher limit value, Ellenberg N shows most deviations from the reference value towards the lower limit value (see N in the respective plot headers). However, the deviations are largely small and within the limits for good ecological condition (0.6 - 1).


[Klikk here](tilrettelegge.html) to see the next step in the analysis, aggregation and standardization of the indicator data set.


