---
title: "Fremmede arter på fjellet"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    
knit: (function(input_file, encoding) {
  out_dir <- '../docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'fremmede-arter-fjell.html'))})
---

```{r setup, include=FALSE, message=FALSE}
library(knitr)
library(tibble)
library(dplyr)
library(stringr)
library(tidyr)
library(tidyverse)
library(ggplot2)
library(hrbrthemes)
library(viridis)
knitr::opts_chunk$set(echo = TRUE)
```

## Beskrivelse
Fremmede arter kan endre økologiske funskjoner og utkonkurrere hjemlige arter. Det er imidlertidig få fremmede arter på fjellet i dag. Her viser vi først kunnskapsstatus for påvirkningsfaktoren fremmedartsintroduksjoner i fjelløkosystemer fra Fremmedartslisten 2018, og så evaluerer vi data fra overvåkingsprogrammet ANO fra årene 209-2021 for fremmede plantearter.

## Data import & handling
This document includes two data sets:

Alien species introductions form Fremmedartslista 2018 (pressure data)  
ANO data 2018-2021 (ecological condition data)  
- geography  
- locations 

# Pressure indicator 'Introduction of alien species'

```{r data import fremmedartslista, include = FALSE}

spr <- read.delim('../data/alien species/spr.txt',header=T, sep = '\t')
sp <- read.delim('../data/alien species/sp.txt',header=T, sep = '\t')

```

```{r data handling fremmedartslista, include = FALSE}
#### Data handling

### Fremmedartsliste data
# removing entries for Svalbard
sprN <- subset(spr, grepl('N', Id))
spN <- subset(sp, grepl('N', Id))
# checking how many colonizations via the main intro categories
sprN$num=1
for (i in unique(sprN$mcat) ) {
  
  print(i)
  print(sum(sprN[sprN$mcat==i,'num']))
}

# merging sprN & spN
colnames(spN)
spN <- spN[,c(2:7,10:13,16,19,24:28,31:35,42:43,53,56)]
sprspN <- merge(sprN, spN, by='TaxonId', all=T)
head(sprspN)
summary(sprspN)
sprspN[sprspN$TaxonId==135279,'C_koloniserte_naturtyper']

# splitting the NiN information from one col for all NiN-types to one col for each NiN-type
split_into_multiple <- function(column, pattern = ", ", into_prefix){
  cols <- str_split_fixed(column, pattern, n = Inf)
  # Sub out the ""'s returned by filling the matrix to the right, with NAs which are useful
  cols[which(cols == "")] <- NA
  cols <- as_tibble(cols)
  # name the 'cols' tibble as 'into_prefix_1', 'into_prefix_2', ..., 'into_prefix_m' 
  # where m = # columns of 'cols'
  m <- dim(cols)[2]
  
  names(cols) <- paste(into_prefix, 1:m, sep = "_")
  return(cols)
}

sprspN.x <- sprspN %>% 
  bind_cols(split_into_multiple(.$C_koloniserte_naturtyper, ",", "NiN")) #%>% 
# selecting those that start with 'NiN_' will remove the original 'type' column
# select(attr, starts_with("NiN_"))

# tidying up the dataframe by gathering all the NiN-cols into one column
sprspN.x <- gather(sprspN.x, 'no', 'NiN', -(1:37) )
colnames(sprspN.x)
dim(sprspN.x)
# getting rid of the comment-column (11), and the redundant NiN-columns ("C_koloniserte_naturtyper" & "no")
sprspN.x <- sprspN.x[,c(1:10,12:36,39)]

# removing duplicate rows (arising from NiN-columns with NA's)
sprspN.x <- distinct(sprspN.x)
dim(sprspN.x)

# splitting the information on 'Arten_finnes_i_foelgende_fylker_omraader' from one col for all fylker to one col for each fylke
sprspN.xx <- sprspN.x %>% 
  bind_cols(split_into_multiple(.$Arten_finnes_i_foelgende_fylker_omraader, ",", "fylke")) #%>% 
# selecting those that start with 'fylke_' will remove the original 'type' column
# select(attr, starts_with("fylke_"))

# tidying up the dataframe by gathering all the fylke-cols into one column
sprspN.xx <- gather(sprspN.xx, 'no', 'fylke', -(1:36) )
colnames(sprspN.xx)
dim(sprspN.xx)
# getting rid of the redundant fylke-columns ("Arten_finnes_i_foelgende_fylker_omraader" & "no")
sprspN.xx <- sprspN.xx[,c(1:31,33:36,38)]

# removing duplicate rows (arising from fylke-columns with NA's)
sprspN.xx <- distinct(sprspN.xx)
dim(sprspN.xx)

# there is codes from NiN1.0 (livsmedium) and NiN2.0 (habitat types)
sprspN.xx$NiNver <- 'NiN2'
sprspN.xx[grepl('S', sprspN.xx$NiN),'NiNver'] <- 'NiN1'
sprspN.xx$NiNver <- as.factor(sprspN.xx$NiNver)
summary(sprspN.xx)

# harmonizing varying NiN-levels (grunntype, hovedtype, kartleggingsenhet) within NiN2.0 to hovedtype
sprspN.xx[sprspN.xx$NiNver=='NiN2','NiN'] <- gsub("\\-.*","",sprspN.xx[sprspN.xx$NiNver=='NiN2','NiN']) # removing basic type and mapping unit suffixes like '-4' and '-C-4'
sprspN.xx$NiN <- as.factor(sprspN.xx$NiN)

# creating variable for NiN-hovedtypegruppe from the NiN-column
sprspN.xx$NiNH <- NA
sprspN.xx[sprspN.xx$NiNver=='NiN2','NiNH'] <- substr(sprspN.xx[sprspN.xx$NiNver=='NiN2','NiN'],1,1)
sprspN.xx$NiNH <- as.factor(sprspN.xx$NiNH)
summary(sprspN.xx)

levels(sprspN.xx$Seksuell_reproduksjon)
levels(sprspN.xx$Seksuell_reproduksjon) <- c('nei','ja')
sprspN.xx[is.na(sprspN.xx$Seksuell_reproduksjon),'Seksuell_reproduksjon'] <- 'nei'
levels(sprspN.xx$Aseksuell_reproduksjon)
levels(sprspN.xx$Aseksuell_reproduksjon) <- c('nei','ja')
sprspN.xx[is.na(sprspN.xx$Seksuell_reproduksjon),'Seksuell_reproduksjon'] <- 'nei'

sprspN.xx$repro <- as.factor(paste(sprspN.xx$Seksuell_reproduksjon,sprspN.xx$Aseksuell_reproduksjon, sep='_'))

levels(sprspN.xx$repro) <- c('both','sexual','asexual','unknown','unknown')

# tidying up the workspace
datN <- sprspN.xx
rm(list=ls(pattern="sp"))

# removing all NiN1 rows for the moment
datn <- datN[datN$NiNver=='NiN2',]
datn$NiN <- factor(datn$NiN)
datn$num=1
### only taking into account species assessed as alien
datn <- datn[datn$Vurderingskategori=='AlienSpecie',]

### creating a column for regions
datn$fylke <- as.factor(datn$fylke)
datn$region <- datn$fylke
levels(datn$fylke)
levels(datn$region) <- c('O','N','O','V','Ja','M','N','M','O','O','V','V','M','Sv','S','N','S','O','O','S')
# removing the fylke variable and the resulting duplicate rows
dim(datn)
datn$fylke <- NULL
dim(datn)
datn <- distinct(datn)
dim(datn)

### only taking into account invasions into mountain ecosystems
head(datn)
unique(datn$NiN)
datn.fjell <- datn[datn$NiN=='T3' | datn$NiN=='T7' | datn$NiN=='T14' | datn$NiN=='T22'| datn$NiN=='T9'| datn$NiN=='T10'| datn$NiN=='T26',]
# merging the mountain ecosystem types to one main ecosytem type (fjell) & removing duplicate rows (due to mountain types merged)
datn.fjell$NiN <- factor(datn.fjell$NiN)
datn.fjell$NiNx <- datn.fjell$NiN
levels(datn.fjell$NiN) <- c('fjell','fjell','fjell','fjell')
dim(datn.fjell)
datn.fjell <- distinct(datn.fjell)
dim(datn.fjell)
# removing the muskoxen, since they are re-introduced
datn.fjell <- datn.fjell[datn.fjell$navn!='moskusfe' & !is.na(datn.fjell$navn),]

# creating one df for the country total and removing duplicate rows from merging regions to country
datn.fjell.T <- datn.fjell
dim(datn.fjell.T)
datn.fjell.T$region <- NULL
dim(datn.fjell.T)
datn.fjell.T <- distinct(datn.fjell.T)
dim(datn.fjell.T)



```

We can plot the cumulative introduction of alien species over time for each region
``` {r alien introduction result, echo = FALSE, fig.show='hide'}

# summarizing
h.T <- hist(datn.fjell.T$Foerst_observert2,xlim=c(1730,2030),breaks=30,main='Norge total',xlab='',ylab='')
h.N <- hist(datn.fjell[datn.fjell$region=='N','Foerst_observert2'],xlim=c(1730,2030),ylim=c(0,110),breaks=30,main='Nord-Norge',xlab='',ylab='')
h.M <- hist(datn.fjell[datn.fjell$region=='M','Foerst_observert2'],xlim=c(1730,2030),ylim=c(0,110),breaks=30,main='Midt-Norge',xlab='',ylab='')
h.V <- hist(datn.fjell[datn.fjell$region=='V','Foerst_observert2'],xlim=c(1730,2030),ylim=c(0,110),breaks=30,main='Vestlandet',xlab='',ylab='')
h.O <- hist(datn.fjell[datn.fjell$region=='O','Foerst_observert2'],xlim=c(1730,2030),ylim=c(0,110),breaks=30,main='Østlandet',xlab='',ylab='')
h.S <- hist(datn.fjell[datn.fjell$region=='S','Foerst_observert2'],xlim=c(1730,2030),ylim=c(0,110),breaks=30,main='Sørlandet',xlab='',ylab='')

# one figure for all
cols <- c("dark grey", "#FFB25B", "#2DCCD3","#004F71", "#7A9A01", "#93328E") 
```

``` {r alien introduction plot, echo = FALSE}
par(mfrow=c(1,1),mar=c(5,5.5,2,1))
with( h.T, plot(mids,cumsum(counts),type='l',col= cols[1],lwd=3,
                ylim=c(0,20), frame=F,main='',yaxt='n',
                xlab='?rstall',ylab='Antall fremmedartsintroduksjoner \n i fjell (kumulativt)') )
axis(side=2, labels= c('0','','10','','20'),at=seq(0,20,5))
with(h.M, points(mids,cumsum(counts),type='l',col= cols[5],lwd=3) )
with(h.V, points(mids,cumsum(counts),type='l',col= cols[4],lwd=3) )
with(h.O, points(mids,cumsum(counts),type='l',col= cols[2],lwd=3) )
with(h.S, points(mids-0.5,cumsum(counts),type='l',col= cols[3],lwd=3) )
with(h.N, points(mids+0.5,cumsum(counts),type='l',col= cols[6],lwd=3) )
legend('topleft',legend=c('Norge total','Nord-Norge','Midt-Norge','Vestlandet','Østlandet','Sørlandet'), 
       col=cols[c(1,6,5,4,2,3)],
       lty=1,lwd=3,bty='n')
```

The introductions cover 17 species within 2 kingdoms - all 11 animal species are birds  
```{r, echo = FALSE}
data.frame(Plantae=nrow(datn.fjell.T[datn.fjell.T$kingdom=='Plantae',]),
           Fungi=nrow(datn.fjell.T[datn.fjell.T$kingdom=='Fungi',]),
           Chromista=nrow(datn.fjell.T[datn.fjell.T$kingdom=='Chromista',]),
           Animalia=nrow(datn.fjell.T[datn.fjell.T$kingdom=='Animalia',]),
           Aves=nrow(datn.fjell.T[datn.fjell.T$class=='Aves',])
)
```

The introductions cover 3 ecosystem types  
```{r, echo = FALSE}
data.frame(Fjellhei_leside_tundra=nrow(datn.fjell.T[datn.fjell.T$NiNx=='T3',]),
           Snøleie=nrow(datn.fjell.T[datn.fjell.T$NiNx=='T7',]),
           Rabbe=nrow(datn.fjell.T[datn.fjell.T$NiNx=='T14',]),
           Fjellgrashei=nrow(datn.fjell.T[datn.fjell.T$NiNx=='T22',])
)
```

# Condition indicator 'Area without cover of alien plant species'

This indicator is defined as 'Area with alien species cover' for plant species since systematically collected data on alien species are available for plants.

```{r data import ANO, include = FALSE}

ANOpkt <- read.table(file="../data/varmekrav/PunktData_all.txt", dec=',', sep = '\t', header=T)
ANOpkt.geo <- read.table(file="../data/varmekrav/ANO_geo.csv", sep = ',', header=T)

```

```{r data handling ANO, include = FALSE}
#### Data handling


### preparing ANO punkt-data and merging with geographical imformation
head(ANOpkt)

colnames(ANOpkt)
# simplfy punkt data
pkt.vars <- c("GlobalID", "ANO_flate_ID", "ANO_punkt_ID", 
              "Hovedoekosystem_rute", "Hovedoekosystem_sirkel", "Andel_hovedoekosystem_sirkel", 
              "Hovedtype_rute", "Kartleggingsenhet_rute", 
              "Hovedtype_sirkel", "Kartleggingsenhet_sirkel", "Andel_Kartleggingsenhet_sirkel",
              "Arter_tilstede", "Dekning_fremmedarter")

ANOpkt <- ANOpkt[,pkt.vars]
ANOpkt <- ANOpkt[!is.na(ANOpkt$GlobalID),]

# merging geographical information with ANOpkt
# removing duplicate ANOID's (due to multiple historical municipalities)
ANOpkt.geo <- ANOpkt.geo[!duplicated(ANOpkt.geo$ANOID), ]

head(ANOpkt.geo)
# fixing names and factor levels in ANOpkt.geo
colnames(ANOpkt.geo)[c(12,16,21)] <- c("Year","Region","Region_no")
ANOpkt.geo$CountyName <- as.factor(ANOpkt.geo$CountyName)
levels(ANOpkt.geo$CountyName) <- c("Oestfold","Akershus",         
                                    "Aust-Agder","Buskerud",         
                                    "Finnmark","Hedmark",          
                                    "Hordaland","Moere og Romsdal",
                                    "Nordland","Oppland",          
                                    "Oslo","Rogaland",         
                                    "Sogn og Fjordane","Telemark",         
                                    "Troendelag","Troms",          
                                    "Vest-Agder","Vestfold") 
ANOpkt.geo$Fylke <- as.factor(ANOpkt.geo$Fylke)
levels(ANOpkt.geo$Fylke) <- c("Oestfold","Akershus",         
                                   "Aust-Agder","Buskerud",         
                                   "Finnmark","Hedmark",          
                                   "Hordaland","Moere og Romsdal",
                              "Nord-Troendelag",
                                   "Nordland","Oppland",          
                                   "Oslo","Rogaland", 
                              "Hordaland","Moere og Romsdal",
                              "Soer-Troendelag",
                                   "Sogn og Fjordane","Telemark",         
                                   "Troendelag","Troms",          
                                   "Vest-Agder","Vestfold") 
# trimming ANOpkt.geo
colnames(ANOpkt.geo)
ANOpkt.geo <- ANOpkt.geo[,c(4,8,9,12,15,16)]
head(ANOpkt.geo)

colnames(ANOpkt)
colnames(ANOpkt.geo)
ANOpkt <- merge(x=ANOpkt, y= ANOpkt.geo,by.x="ANO_flate_ID", by.y="ANOID", all.x=T)
head(ANOpkt)


```

We define the reference value to reflect the abscence of alien species (i.e. 100% cover without alien species), the limit value for good ecological condition is set to 95% (expert-opinion based), and the minimum value is 0.  
``` {r defining the reference condition}

alien.ref <- data.frame(ref=100,lim=95,min=0)
```

Next we scale the alien species cover data against the threshold values in alien.ref.  
``` {r scaling, echo = FALSE, include=FALSE}

#### scaled values ####
r.s <- 1    # reference value
l.s <- 0.6  # limit value
a.s <- 0    # abscence of indicator, or indicator at maximum

#### function for calculating scaled values for measured value ####

## scaling function including truncation
scal <- function() {
  # place to hold the result
  x <- numeric()
  if (maxmin < ref) {
    # values >= the reference value equal 1
    if (val >= ref) {x <- 1}
    # values < the reference value and >= the limit value can be deducted from the linear relationship between these two
    if (val < ref & val >= lim) {x <- (l.s + (val-lim) * ( (r.s-l.s) / (ref-lim) ) )}
    # values < the limit value and > maxmin can be deducted from the linear relationship between these two
    if (val < lim & val > maxmin) {x <- (a.s + (val-maxmin) * ( (l.s-a.s) / (lim-maxmin) ) )}
    # value equals or lower than maxmin
    if (val <= maxmin) {x <-0}
  } else {
    # values <= the reference value equal 1
    if (val <= ref) {x <- 1}
    # values > the reference value and <= the limit value can be deducted from the linear relationship between these two
    if (val > ref & val <= lim) {x <- ( r.s - ( (r.s - l.s) * (val - ref) / (lim - ref) ) )}
    # values > the limit value and < maxmin can be deducted from the linear relationship between these two
    if (val > lim) {x <- ( l.s - (l.s * (val - lim) / (maxmin - lim) ) )}
    # value equals or larger than maxmin
    if (val >= maxmin) {x <-0}
  }
  return(x)
  
}



#### creating dataframe to holde the results ####
ANOpkt.mount <- ANOpkt[ANOpkt$Hovedoekosystem_rute=='fjell' | ANOpkt$Hovedoekosystem_sirkel=='fjell',]
ANOpkt.mount$Hovedtype_rute <- factor(ANOpkt.mount$Hovedtype_rute)
ANOpkt.mount$Kartleggingsenhet_rute <- factor(ANOpkt.mount$Kartleggingsenhet_rute)

results.mount.alien <- list()
# choose columns for site description
colnames(ANOpkt.mount)
results.mount.alien[['original']] <- ANOpkt.mount[,c(2,1,3,4:18)]

# add column for the indicators
results.mount.alien[['original']]$alien <- as.numeric(NA)

summary(results.mount.alien[['original']])
results.mount.alien[['original']]$Region <- as.factor(results.mount.alien[['original']]$Region)
results.mount.alien[['original']]$GlobalID <- as.factor(results.mount.alien[['original']]$GlobalID)
results.mount.alien[['original']]$ANO_flate_ID <- as.factor(results.mount.alien[['original']]$ANO_flate_ID)
results.mount.alien[['original']]$ANO_punkt_ID <- as.factor(results.mount.alien[['original']]$ANO_punkt_ID)
results.mount.alien[['original']]$Hovedoekosystem_sirkel <- as.factor(results.mount.alien[['original']]$Hovedoekosystem_sirkel)
results.mount.alien[['original']]$Hovedoekosystem_rute  <- as.factor(results.mount.alien[['original']]$Hovedoekosystem_rute )
results.mount.alien[['original']]$Kartleggingsenhet_sirkel <- as.factor(results.mount.alien[['original']]$Kartleggingsenhet_sirkel)
results.mount.alien[['original']]$Hovedtype_sirkel    <- as.factor(results.mount.alien[['original']]$Hovedtype_sirkel)


# roll out
results.mount.alien[['scaled']] <- results.mount.alien[['original']]


#### calculating scaled and non-truncated values for the indicators based on the dataset ####
for (i in 1:nrow(ANOpkt.mount) ) {  #
  tryCatch({
    print(i)
    print(paste(ANOpkt.mount$ANO_flate_ID[i]))
    print(paste(ANOpkt.mount$ANO_punkt_ID[i]))
#    ANOpkt.mount$Hovedoekosystem_sirkel[i]
#    ANOpkt.mount$Hovedoekosystem_rute[i]


    if (ANOpkt.mount$Hovedoekosystem_sirkel[i]=='fjell' & !is.na(ANOpkt.mount$Dekning_fremmedarter[i]) ) {
      
      # area without alien species      
        val <- 100-ANOpkt.mount$Dekning_fremmedarter[i]
        ref <- alien.ref[,'ref']
        lim <- alien.ref[,'lim']
        maxmin <- alien.ref[,'min']
        # coercing x into results.mount.alien dataframes in list
        results.mount.alien[['scaled']][i,'alien'] <- scal() 
        results.mount.alien[['original']][i,'alien'] <- val 
    }
    
  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}

```

Here is the result for Norway  
``` {r result}

# overall result
summary(results.mount.alien[['scaled']][,'alien'])
```

and by region (C = Central, E = Eastern, N = Northern, S = Southern, W = Western)  
``` {r region result, echo=F, warning = F, message = F}

# sample size
sample_size = results.mount.alien[['scaled']] %>% group_by(Region) %>% summarize(num=n())

# Plot
results.mount.alien[['scaled']] %>%
  left_join(sample_size) %>%
  mutate(myaxis = paste0(Region, "\n", "n=", num)) %>%
  ggplot( aes(x=myaxis, y=alien, fill=Region)) +
    geom_violin(width=1.4) +
#    geom_boxplot(width=0.1, color="grey", alpha=0.2) +
    scale_fill_viridis(discrete = TRUE) +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=11)
    ) +
    ggtitle("Scaled index result") +
    xlab("Region") +
  ylab("Alien species index")

```

Most indices are equal or close to 1 (optimal). Serious deviations do occur in Eastern and Northern Norway, but they are very seldom. Southern and Western Norway do not have a single index value below 1, i.e. there were no alien species registrations at all.  


[Klikk here](tilrettelegge.html) to see the next step in the analysis, aggregation and standardization of the indicator data set.

