---
title: "Påvirkningsfigurer"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    
knit: (function(input_file, encoding) {
  out_dir <- '../docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'paavirkninger.html'))})
---

```{r setup, include=FALSE, message=FALSE}
library(knitr)
library(DT)
library(dplyr)
library(plyr)
library(matrixStats)
knitr::opts_chunk$set(echo = TRUE)
```

Her reger jeg ut og plotter de aggregerte tilstandsverdiene for hver generelle påvirkningskategori. Skriptet følger etter det skriptet som heter [plotting.R](plotting.html).

Skriptet er modifisert fra det som Simon Jakobsson skrev for skograpporten i 2021.

# Step 1: Basics & load data

Indicator and weights files from aggregation script
```{r}
allind  <- read.csv("../output/allind_temp.csv", row.names=1)
weights <- read.csv("../output/weights_temp.csv", row.names=1)
```

Number of simulations to run
```{r}
nsim <- 100 # bør kjøres på 10 000 til slutt
```

Total number of indicators
```{r}
(nind <- length(unique(allind$X)))
```

Short names for ecosystem characteristics
```{r}
pressures <- c("are", "kli", "fef", "bes", "fre")
```

```{r}
pressures2 <- c("Arealinngrep", "Klima", "Forurensing",
                     "Beskatning", "Fremmede arter")
```


Empty indicator <-> pressures matrix
```{r}
ind_press <- data.frame(matrix(nrow=nind, ncol=6))
names(ind_press) <- c("IND", pressures)
ind_press$IND <- unique(allind$X)
```

```{r}
ind_press$are <- 
  ifelse(
    ind_press$IND=="alien"|
    ind_press$IND=="breareal"|
    ind_press$IND=="EllLHigh"|
    ind_press$IND=="EllLow"|
    #ind_press$IND=="EllNHigh"|
    #ind_press$IND=="EllNLow"|
    ind_press$IND=="fjellindeks"|
    ind_press$IND=="fjellrev"|
    #ind_press$IND=="fjellrype"|
    ind_press$IND=="INON"|
    ind_press$IND=="jerv"|
    ind_press$IND=="kongeørn"|
    ind_press$IND=="ndvi"
    #ind_press$IND=="snodekke"|
    #ind_press$IND=="snodybde"|
    #ind_press$IND=="smaagnagere"|
    #ind_press$IND=="varmekrav"|
    #ind_press$IND=="vinterregn"
    ,
    1,0)


ind_press$kli  <- 
  ifelse(
    ind_press$IND=="alien"|
    ind_press$IND=="breareal"|
    ind_press$IND=="EllLHigh"|
    ind_press$IND=="EllLow"|
    ind_press$IND=="EllNHigh"|
    ind_press$IND=="EllNLow"|
    ind_press$IND=="fjellindeks"|
    ind_press$IND=="fjellrev"|
    ind_press$IND=="fjellrype"|
    #ind_press$IND=="INON"|
    #ind_press$IND=="jerv"|
    ind_press$IND=="kongeørn"|
    ind_press$IND=="ndvi"|
    ind_press$IND=="snodekke"|
    ind_press$IND=="snodybde"|
    ind_press$IND=="smaagnagere"|
    ind_press$IND=="varmekrav"|
    ind_press$IND=="vinterregn"
    ,
    1,0)
ind_press$fef  <- 
  ifelse(
    #ind_press$IND=="alien"|
    #ind_press$IND=="breareal"|
    #ind_press$IND=="EllLHigh"|
    #ind_press$IND=="EllLow"|
    #ind_press$IND=="EllNHigh"|
    ind_press$IND=="EllNLow"|
    #ind_press$IND=="fjellindeks"|
    #ind_press$IND=="fjellrev"|
    #ind_press$IND=="fjellrype"|
    #ind_press$IND=="INON"|
    #ind_press$IND=="jerv"|
    #ind_press$IND=="kongeørn"|
    ind_press$IND=="ndvi"
    #ind_press$IND=="snodekke"|
    #ind_press$IND=="snodybde"|
    #ind_press$IND=="smaagnagere"|
    #ind_press$IND=="varmekrav"|
    #ind_press$IND=="vinterregn"
    ,
    1,0)


ind_press$bes <- 
  ifelse(
    #ind_press$IND=="alien"|
    #ind_press$IND=="breareal"|
    #ind_press$IND=="EllLHigh"|
    #ind_press$IND=="EllLow"|
    #ind_press$IND=="EllNHigh"|
    #ind_press$IND=="EllNLow"|
    ind_press$IND=="fjellindeks"|
    ind_press$IND=="fjellrev"|
    ind_press$IND=="fjellrype"|
    #ind_press$IND=="INON"|
    ind_press$IND=="jerv"
    #ind_press$IND=="kongeørn"|
    #ind_press$IND=="ndvi"|
    #ind_press$IND=="snodekke"|
    #ind_press$IND=="snodybde"|
    #ind_press$IND=="smaagnagere"|
    #ind_press$IND=="varmekrav"|
    #ind_press$IND=="vinterregn"
    ,
    1,0)

ind_press$fre <- ifelse(
    ind_press$IND=="alien"
    ,
    1,0)
```


```{r, eval=F}
write.csv(ind_press, "../output/ind_press.csv")
```

# Step 2: Aggregated pressure calculations

Loop factors
```{r}
#periods <- unlist(unique(allind$a_period))
regions <- unlist(unique(allind$reg))
```
Empty aggregation data frame
```{r}
agg_ind <- data.frame()
w_boot <- data.frame()
```

```{r}
for (n in 1:length(pressures)){
  
  # Select by ind_press data frame
  temp_press <- 
    cbind(
      ind_press$IND, 
      ind_press[names(ind_press)%in%pressures[n]])
  
  temp_press <- temp_press[temp_press[,2]==1,]
  
  temp_ind <- allind[allind$X %in% temp_press$`ind_press$IND`,]


  for (k in 1:length(regions)){
      
    print(paste(pressures[n], regions[k]))
# Subset region
    temp_ind3 <- subset(temp_ind, reg==regions[k])
    #temp_ind3 <- data.frame(temp_ind3[,-(1:3)])
    #names(temp_ind3) <- names(temp_ind2)[-(1:3)]
      
# weights
    temp_weights <- weights[weights$reg==regions[k],]
    temp_weights <- temp_weights[temp_weights$X %in%
                                   unique(temp_ind3$X),]
    
# Empty sample mean vector
    temp_x_boot <- NULL
# Empty sample values vector  
    temp_x <- NULL
    temp_x2 <- NULL
        
 for (i in 1:length(unique(temp_ind3$X))) {
          
          print(paste(
            pressures[n], 
            regions[k],
            unique(temp_ind3$X)[i]))
          
# Sample from indicator i
    start <- length(temp_x)
   
    temp_x[(start+1):(start+nsim)] <- 
      tryCatch(
      sample(
       temp_ind3$val[temp_ind3$X ==
                       unique(temp_ind3$X)[i]], 
                        nsim, replace = T), 
                        error=function(e){})
          
    temp_x2[(start+1):(start+nsim)] <- unique(temp_ind3$X)[i]
        } 
        
    temp_x3 <- as.data.frame(cbind(val=temp_x, 
                                      ind =temp_x2))
    setDT(temp_x3)
    temp_x3$sim <- rep(1:nsim, 
                       times=length(unique(temp_ind3$X)))
    temp_x3$val <- as.numeric(temp_x3$val)
    temp_x3 <- data.table::dcast(temp_x3,
                                    sim~ind,
                                    value.var="val")
       
    mat <- as.matrix(temp_x3[,-1])
    x_boot <- rowWeightedMeans(mat, temp_weights$val)
      
# Info data
    pres <- pressures[n]
    reg <- regions[k]
      
    w_boot <- rbind(w_boot, data.frame(pres, reg, x_boot))
      
# Quantiles
    temp_Q <- tryCatch(quantile(
      x_boot,c(0.025, 0.5, 0.975)), 
      error=function(e){})
      
      # Compile data
    agg_ind <- rbind(
      agg_ind, 
      data.frame(pres, 
                 reg,
                 ifelse(is.null(temp_Q), NA, temp_Q[1]),
                 ifelse(is.null(temp_Q), NA, temp_Q[2]),
                 ifelse(is.null(temp_Q), NA, temp_Q[3]))) 
    
  }
}

# Rename columns
names(agg_ind) <- c("ind", "reg",  "low", "med", "upp")
```

## Names
```{r}
temp_names <- data.frame(names(ind_press[,-1]))
names(temp_names) <- "work"
temp_names$plot <- pressures2
```


# Step 9: Plotting separate     

```{r}

for (j in 1:length(regions)){

  TEMP_Q <- agg_ind[
    agg_ind$reg==regions[j],]
    
  TEMP_Q$ind_2 <- c(1:nrow(TEMP_Q))
  
  TEMP_Q$ind <- factor(temp_names$plot[match(TEMP_Q$ind, temp_names$work)])
  
    
    temp_per <- 2021
    
    write.csv(TEMP_Q, paste("../output/tables/TEMP_Q_", regions[j], "_", temp_per, ".csv", sep=""))
    
    png(paste("../output/aggregated_plots/pressure plots/region_", regions[j],"_", ".png", sep=""), units="in", width=4, height=11, res=600)
    
    par(xaxt="n", mfrow=c(1,1), par(mar=c(25,5,1,1)))
    plot(TEMP_Q$med~TEMP_Q$ind_2, ylab="Skalerte indikatorverdier", xlab="",
         type="n", ylim=c(0,1.1), xlim=c(0.5,5.5), cex.lab=1.2)
    abline(h=1, col="blue", lwd=2)
    abline(h=0.6, col="red", lwd=2, lty=2)
    
    # add values and CI's
  arrs <- data.frame(
           x  = TEMP_Q$ind_2[1:nrow(TEMP_Q)],
           y = TEMP_Q$med[1:nrow(TEMP_Q)],
           high = TEMP_Q$upp[1:nrow(TEMP_Q)],
           low = TEMP_Q$low[1:nrow(TEMP_Q)]
  )   
  arrs$diff <- arrs$y-arrs$high
  arrs$diff2 <- arrs$y-arrs$low
  arrs2 <- arrs
  arrs <- arrs[arrs$diff != 0,]    
  arrs <- arrs[arrs$diff2 != 0,]    
  
  arrows(arrs$x, arrs$y,
        arrs$x, arrs$high, 
            angle=90, length=0.05, col="grey")
    
  arrows(arrs$x, arrs$y,
        arrs$x, arrs$low, 
            angle=90, length=0.05, col="grey")
    
  points(
    arrs2$x, arrs2$y, 
           pch=21, bg="dark grey", cex=1.5)
    
    # Name axes #
    lablist.x<-as.vector(TEMP_Q$ind)
    
    axis(1, at=seq(1,  nrow(TEMP_Q), by=1), labels = FALSE)
    
    # Change first command number to align labels
    text(TEMP_Q$ind_2+0.2, par("usr")[3]-0.02, labels = lablist.x,pos=2,srt = 90,  xpd = TRUE, cex=1.2)
    
    dev.off()
  
}

```

# Step 10: Plotting combined 

```{r}
regions <- data.frame(ord = c(5,6,4,3,2,1))
regions$navn <- unlist(unique(allind$reg))
regions <- regions[order(regions$ord),]
regions$colours <- c("dark grey", "#FFB25B", "#2DCCD3","#004F71", "#7A9A01", "#93328E")

move <- 4
div <- 8

  
  TEMP_Q <- agg_ind[
    agg_ind$reg==regions$navn[2],]
  
  TEMP_Q$ind_2 <- c(1:nrow(TEMP_Q))
  TEMP_Q$ind <- factor(temp_names$plot[match(TEMP_Q$ind, temp_names$work)])
  
  temp_per <- 2021
  
  write.csv(TEMP_Q, paste("../output/tables/TEMP_Q_comb_", regions$navn[2], "_", temp_per, ".csv", sep=""))
  

  png(paste("../output/aggregated_plots/pressure plots/comb_", temp_per, ".png", sep=""), units="in", width=6, height=11, res=600)
  
  par(xaxt="n", mfrow=c(1,1), par(mar=c(25,5,1,1)))
  plot(TEMP_Q$med~TEMP_Q$ind_2, ylab="Skalerte indikatorverdier", xlab="",
       type="n", ylim=c(0,1.1), xlim=c(0.5,5.5), cex.lab=1.2)
  abline(h=1, col="blue", lwd=2)
  abline(h=0.6, col="red", lwd=2, lty=2)
  
  # add values and CI's
  arrs <- data.frame(
           x  = TEMP_Q$ind_2[1:nrow(TEMP_Q)],
           y = TEMP_Q$med[1:nrow(TEMP_Q)],
           high = TEMP_Q$upp[1:nrow(TEMP_Q)],
           low = TEMP_Q$low[1:nrow(TEMP_Q)]
  )   
  arrs$diff <- arrs$y-arrs$high
  arrs$diff2 <- arrs$y-arrs$low
  arrs2 <- arrs
  arrs <- arrs[arrs$diff != 0,]    
  arrs <- arrs[arrs$diff2 != 0,] 
  
  arrows(arrs$x+(2-move)/div,
         arrs$y,
         arrs$x+(2-move)/div,
         arrs$high, 
          angle=90, length=0.05, col=regions$colours[2])
  
  arrows(arrs$x+(2-move)/div,
         arrs$y,
         arrs$x+(2-move)/div,
         arrs$low, 
          angle=90, length=0.05, col=regions$colours[2])
  
  points(arrs2$x+(2-move)/div,
         arrs2$y, 
         pch=21, bg=regions$colours[2], cex=1.5)
  
  # Name axes #
  lablist.x<-as.vector(TEMP_Q$ind)
  
  axis(1, at=seq(1,  nrow(TEMP_Q), by=1), labels = FALSE)
  
  # Change first command number to align labels
  text(TEMP_Q$ind_2+0.2, par("usr")[3]-0.02, labels = lablist.x,pos=2,srt = 90,  xpd = TRUE, cex=1.2)
  
  
  for (j in 3:length(regions$navn)){
    
    TEMP_Q <- agg_ind[
      agg_ind$reg==regions$navn[j],]
    
    TEMP_Q$ind_2 <- c(1:nrow(TEMP_Q))
    TEMP_Q$ind <- factor(
      temp_names$plot[match(TEMP_Q$ind, temp_names$work)])
    
    temp_per <- 2021
    
    write.csv(TEMP_Q, paste("../output/tables/TEMP_Q_comb_", regions$navn[j], "_", temp_per, ".csv", sep=""))
    
    
    # add values and CI's
    arrs <- data.frame(
           x  = TEMP_Q$ind_2[1:nrow(TEMP_Q)],
           y = TEMP_Q$med[1:nrow(TEMP_Q)],
           high = TEMP_Q$upp[1:nrow(TEMP_Q)],
           low = TEMP_Q$low[1:nrow(TEMP_Q)]
  )   
  arrs$diff <- arrs$y-arrs$high
  arrs$diff2 <- arrs$y-arrs$low
  arrs2 <- arrs
  arrs <- arrs[arrs$diff != 0,]    
  arrs <- arrs[arrs$diff2 != 0,] 
  
  arrows(arrs$x+(j-move)/div,
         arrs$y,
         arrs$x+(j-move)/div,
         arrs$high, 
          angle=90, length=0.05, col=regions$colours[j])
  
  arrows(arrs$x+(j-move)/div,
         arrs$y,
         arrs$x+(j-move)/div,
         arrs$low, 
          angle=90, length=0.05, col=regions$colours[j])
  
  points(arrs2$x+(j-move)/div,
         arrs2$y, 
         pch=21, bg=regions$colours[j], cex=1.5)
    
  }    
  
  dev.off()
```

