---
title: "Vegetasjonenens_varmekrav"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true

knit: (function(input_file, encoding) {
  out_dir <- '../docs';
  rmarkdown::render(input_file,
                    encoding=encoding,
                    output_file=file.path(dirname(input_file), out_dir, 'vegetasjonens_varmekrav.html'))})
---
  
```{r setup, include=FALSE, message=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

### R-libraries nedded
```{r setup, message=FALSE, warning = FALSE}
library(knitr)
library(tidyr)
library(stringr)
library(tidyverse)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```

# Analyses of heat rquirement of species in a plant community 
- Data import & handling
- Building the reference
- Scaling of the evaluation data to an ecological-condition index

## Data import & handling
We have three data sources for this indicator:

Heat_requirement indicator data  
NiN generalized species lists for mountain ecosystems  
ANO data  
- geography  
- locations  
- species 

```{r include = FALSE}
setwd("C:/Users/joachim.topper/OneDrive - NINA/work/R projects/github/IBECA")

ind_swe <- read.table(file="data/varmekrav/ind_swe.txt", sep = '\t', header=T, quote = '')

load("data/varmekrav/Eco_State.RData")

ANOpkt <- read.table(file="data/varmekrav/PunktData_all.txt", dec=',', sep = '\t', header=T)
ANOpkt.geo <- read.table(file="data/varmekrav/ANO_geo.csv", sep = ',', header=T)
sp <- read.table(file="data/varmekrav/ArtsData_all.txt", dec=',', sep = '\t', header=T)
```

ind_swe is a data frame with a range of species indicators for Swedish plant species. Literature reference:

Tyler, T., Herbertsson, L., Olofsson, J., & Olsson, P. A. (2021). Ecological indicator and traits values for Swedish vascular plants. Ecological Indicators, 120. doi:10.1016/j.ecolind.2020.106923
```{r eval = FALSE}
head(ind_swe)
```

Eco_State.RData is an R-list containing the generalized species lists from NiN. References:

Halvorsen, R., Skarpaas, O., Bryn, A., Bratli, H., Erikstad, L., Simensen, T., & Lieungh, E. (2020). Towards a systematics of ecodiversity: The EcoSyst framework. Global Ecology and Biogeography, 29(11), 1887-1906. doi:10.1111/geb.13164

Halvorsen, R., Bryn, A., & Erikstad, L. (2015). NiNs systemkjerne – teori, prinsipper og inndelingskriterier. – Natur i Norge, Artikkel 1 (versjon 2.0.0)
```{r eval = FALSE}
str(Eco_State)
```

ANOpkt.geo is a data frame with geograhical information for each ANO site and point
```{r eval = FALSE}
head(ANOpkt.geo)
```

ANOpkt is a data frame with ecologcial information for each ANO-point
```{r eval = FALSE}
head(ANOpkt)
```

sp is a data frame with plant species information (occurence and abundance) for each ANO-point
```{r eval = FALSE}
head(sp)
```


```{r include = FALSE}
#### Data handling
### Indicator data
names(ind_swe)[1] <- 'species'
ind_swe$species <- as.factor(ind_swe$species)
ind_swe <- ind_swe[!is.na(ind_swe$species),]
ind_swe[,'species'] <- word(ind_swe[,'species'], 1,2)
ind_swe <- ind_swe[!duplicated(ind_swe[,'species']),]
ind_swe$species <- as.factor(ind_swe$species)

## NiN data
# transposing abundance data for bootstrapping
all.dat <- t(Eco_State$Concept_Data$Species$Species_Data)
all.dat <- as.data.frame(all.dat)
all.dat$sp <- as.factor(as.vector(Eco_State$Concept_Data$Species$Species_List$species))
# only genus and species name
all.dat$sp <- word(all.dat$sp, 1,2)
all.dat$spgr <- as.factor(as.vector(Eco_State$Concept_Data$Species$Species_List$art.code))
# environment data
env.dat <- Eco_State$Concept_Data$Env$Env_Data

## merging NiN lists and indicator values
all.ind.dat <- merge(all.dat,ind_swe, by.x="sp", by.y="species", all.x=T)
all.ind.dat[all.ind.dat==999] <- NA

## extracting mountain NiN-lists
mount.dat <- all.ind.dat[,c("sp",paste(env.dat$ID[env.dat$Nature_Type=="Mountain " | env.dat$Sub_Type=="Mountain" | env.dat$Sub_Type=="Mountain"]),"Heat_requirement")]
# chosing the relevant species lists and assigning the right NiN names
mount.dat <- mount.dat[,c(1:19,21:26,34:36,40)]
colnames(mount.dat)[2:28] <- c('T14-C1','T14-C2','T3-C3C6','T3-C9C12',
                               'T3-C2C5','T3-C8C11','T3-C1a','T3-C1b',
                               'T3-C4','T3-C7','T3-C10','T3-C13',
                               'T3-C14','T7-C1','T7-C2','T7-C3',
                               'T7-C6','T7-C8','T7-C12','T7-C13C14',
                               'T7-C4','T7-C7C9','T7-C5','T7-C10',
                               'T22-C1C3a','T22-C1C3b','T22-C2C4')

# checking max values per kartleggingsenhet
check.dat <- data.frame(NiN=names(mount.dat)[2:28], HeatMax=NA)
for (i in names(mount.dat)[2:28] ) {
  check.dat[check.dat$NiN==i,'HeatMax'] <- max(mount.dat[!is.na(mount.dat[,i]) & mount.dat[,i]>0,c('Heat_requirement')],na.rm=T)
}
check.dat
# Heat_requirement of 11 in T3-C14 is unreasonably high for a mountain type, checking
mount.dat[!is.na(mount.dat$'T3-C14') & mount.dat$'T3-C14'>0,c('T3-C14','Heat_requirement','sp')]
# Stellaria nemorum, mainly a forest species
mount.dat[mount.dat$sp=='Stellaria nemorum',]
# only present in T3-C14 -> omit Stallaria nemorum
mount.dat <- mount.dat[mount.dat$sp!='Stellaria nemorum',]

# translating the abundance classes into %-cover
coverscale <- data.frame(orig=0:6,
                         cov=c(0,1/32,1/8,3/8,0.6,4/5,1)
)

mount.dat.cov <- mount.dat
for (i in 2:28) {
  mount.dat.cov[,i] <- coverscale[,2][ match(mount.dat[,i], 0:6 ) ]
}


## ANO data

colnames(sp)[5] <- 'Dekning'
# simplify punkt data
pkt.vars <- c("Aar", "GlobalID", "ANO_flate_ID", "ANO_punkt_ID", 
              "Hovedoekosystem_rute", "Hovedoekosystem_sirkel", "Andel_hovedoekosystem_sirkel", 
              "Hovedtype_rute", "Kartleggingsenhet_rute", 
              "Hovedtype_sirkel", "Kartleggingsenhet_sirkel", "Andel_Kartleggingsenhet_sirkel",
              "Arter_tilstede", "Dekning_fremmedarter")

ANOpkt <- ANOpkt[,pkt.vars]
ANOpkt <- ANOpkt[!is.na(ANOpkt$GlobalID),]


# removing duplicate ANOID's (due to multiple historical municipalities)
ANOpkt.geo <- ANOpkt.geo[!duplicated(ANOpkt.geo$ANOID), ]

# fixing names and factor levels in ANOpkt.geo
colnames(ANOpkt.geo)[c(12,16,21)] <- c("Year","Region","Region_no")
ANOpkt.geo$CountyName <- as.factor(ANOpkt.geo$CountyName)
levels(ANOpkt.geo$CountyName) <- c("Oestfold","Akershus",         
                                   "Aust-Agder","Buskerud",         
                                   "Finnmark","Hedmark",          
                                   "Hordaland","Moere og Romsdal",
                                   "Nordland","Oppland",          
                                   "Oslo","Rogaland",         
                                   "Sogn og Fjordane","Telemark",         
                                   "Troendelag","Troms",          
                                   "Vest-Agder","Vestfold") 
ANOpkt.geo$Fylke <- as.factor(ANOpkt.geo$Fylke)
levels(ANOpkt.geo$Fylke) <- c("Oestfold","Akershus",         
                              "Aust-Agder","Buskerud",         
                              "Finnmark","Hedmark",          
                              "Hordaland","Moere og Romsdal",
                              "Nord-Troendelag",
                              "Nordland","Oppland",          
                              "Oslo","Rogaland", 
                              "Hordaland","Moere og Romsdal",
                              "Soer-Troendelag",
                              "Sogn og Fjordane","Telemark",         
                              "Troendelag","Troms",          
                              "Vest-Agder","Vestfold") 
# trimming ANOpkt.geo
colnames(ANOpkt.geo)
ANOpkt.geo <- ANOpkt.geo[,c(4,8,9,12,15,16)]

# merging ANOpkt & ANOpkt.geo
ANOpkt <- merge(x=ANOpkt, y= ANOpkt.geo,by.x="ANO_flate_ID", by.y="ANOID", all.x=T)

## add indicator values to species data
# fix species names
head(sp)
sp$Species <- sp$Navn
sp$Species <- sub(".*?_", "", sp$Species) # lose the Norwegian name in the front
sp$Species <- sub("^([^_]*_[^_]*).*", "\\1", sp$Species) # lose subspecies
sp$Species <- str_to_title(sp$Species) # make first letter capital
sp$Species <- gsub("_", " ", sp$Species) # replace underscore with space
sp$Species <- gsub("(-.*)","\\L\\1",sp$Species,perl=TRUE) # make capital letters after hyphon to lowercase

## merge species data with indicators
sp <- merge(x=sp, y= ind_swe,by.x="Species", by.y="species", all.x=T)
summary(sp)

## adding hoved?kosystem, hovetype & kartleggingsenhet
sp <- merge(x=sp, y=ANOpkt[,c("GlobalID","Hovedoekosystem_rute","Hovedtype_rute","Kartleggingsenhet_rute")], 
            by.x="ParentGlobalID", by.y="GlobalID", all.x=T)
# trimming away the points without information on NiN, species or cover
sp <- sp[!is.na(sp$Hovedoekosystem_rute),]
sp <- sp[!is.na(sp$Species),]
sp <- sp[!is.na(sp$Dekning),]


# ...which makes us end up with:  
  
# NiN species lists for mountains matched with indicator values  
# Note that the abundance values in the NiN species lists are maximum abundance!
head(mount.dat.cov)
# Species data from ANO matched with indicator values
head(sp)
```
After some merging and data wrangling we arrive at (i) NiN species lists for mountains matched with Heat-requirement values (Note that the abundance values in the NiN species lists are maximum abundance). From here we can build the reference.
``` {r}
head(mount.dat.cov)
```

...and at (ii) Species data from the ANO monitoring program matched with Heat-requirement values
``` {r}
head(sp)
```

## Building the reference (scaling values)

### concept for the reference
We can calculate cumulative cums of cover for all Heat-requirement levels starting with the highest level. Note that (i) it is maximum cover that is represented in the generalized species lists from NiN, and (ii) we use relative cover, i.e. every species' cover is devided by the total cover of all species summed.
```{r echo = FALSE}
abun.cumsums <- abun.sums <- data.frame(ind=rev(1:9))

for (i in 2:28) {
  df <- mount.dat.cov[,c(29,i)]
  df <- df[!is.na(df[,1]) & df[,2]>0,]
  df[,2] <- df[,2]/sum(df[,2])
  colnames(df) <- c('ind','abun')
  df <- df %>% add_row(ind=1:9, abun=rep(0,9))
  abun.sums[,i] <- rev(with(df,tapply(abun,ind,sum)))
  abun.cumsums[,i] <- cumsum(abun.sums[,i])
}

colnames(abun.cumsums)[2:28] <- colnames(mount.dat.cov)[2:28]
```
Here is an example illustrating how this  cumulative distribution of cover from high to low heat requirement for one ecosystem type in NiN: T3-C4 is a leeside with intermediate access to lime  (Intermediær leside)
```{r echo=FALSE}
plot(abun.cumsums[,1],abun.cumsums[,10],
     type='b',
     xlab="Heat_requirement",
     ylab="Cumulative relative abundance",
     main=colnames(abun.cumsums)[10])
```
Under climate warming we'd expect the plant community to show a shift in dominance structure, where the cover of species with higher heat requirement increases. This may be both due to a cover increase of certain alpine species with higher heat requirement or because of colonization by species having a higher heat requirement than what is represented in the reference condition for Alpine systems.
Under this scenario, we should expect the cumulative sum of relative cover for higher heat requirement to at some point in time exceed the maximum value in the reference distribution.  
Below comes the same plot as above, but now also including the cumulative sum of cover for two ANO-points.

```{r echo = FALSE}
good <- sp[sp$ParentGlobalID=='c0594346-c084-4d61-9c4d-4b0c0d50dd83',c('Species','Dekning','Heat_requirement')]
bad <- sp[sp$ParentGlobalID=='4a39f7af-7aa2-4aaa-9907-e52d08268186',c('Species','Dekning','Heat_requirement')]

abun.cumsums.bad <- abun.sums.bad <- abun.cumsums.good <- abun.sums.good <- data.frame(ind=rev(1:9))

good[,2] <- good[,2]/sum(good[,2])
colnames(good)[2:3] <- c('abun','ind')
good <- good %>% add_row(ind=1:9, abun=rep(0,9))
abun.sums.good[,2] <- rev(with(good,tapply(abun,ind,sum)))
abun.cumsums.good[,2] <- cumsum(abun.sums.good[,2])

bad[,2] <- bad[,2]/sum(bad[,2])
colnames(bad)[2:3] <- c('abun','ind')
bad <- bad %>% add_row(ind=1:9, abun=rep(0,9))
abun.sums.bad[,2] <- rev(with(bad,tapply(abun,ind,sum)))
abun.cumsums.bad[,2] <- cumsum(abun.sums.bad[,2])


plot(abun.cumsums[,1],abun.cumsums[,10],
     type='b',
     xlab="Heat_requirement",
     ylab="cumulative relative abundance",
     main=colnames(abun.cumsums)[10])
points(abun.cumsums.good[,1],abun.cumsums.good[,2],type='b', col='green')
points(abun.cumsums.bad[,1],abun.cumsums.bad[,2],type='b', col='red')
```

The green ANO-point seems to be better aligned with the reference (black), while the red ANO-points obviously have a larger cover of species with higher Heat-requirement.  
All we need to do is now to define at which level of heat requirement the cumulative cover is to be compared between evaluation data and reference? In principal, this should be the heat requirement level before the cumulative curve in the reference reaches its steepest slope. Visual inspection of the plot suggests that to be the case at a Heat_requirement of 3 in the example.   

We evaluate all NiN-types visually and create a table holding this information and with space for the scaling values which are to be defined further down.

```{r echo = FALSE}
heat.scal.tab.boot <- data.frame(NiN=colnames(mount.dat.cov)[2:28],                            ind_level=NA,ref=NA,lim=NA,maxmin=1)
heat.scal.tab.boot$ind_level <- c(3,3,3,3,3,3,3,3,3,3,3,4,4,3,3,3,3,3,3,3,3,3,2,NA,3,3,3)
```

### Definisjon av skaleringsverdier
Next step is to define scaling values for scaling evaluation data to the ecological condition index based on the cumulative cover distributions for each NiN ecosystem type. For this, we bootstrap the NiN species lists, calculate the (relative) cumulative cover for all species at the heat requirement threshold and above in each bootstrap, and then derive limit and reference values from the resulting distribution of bootstraped cumulative cover values.

### Bootstrap function
sp...the species names
abun...the species abundances for every ecosystem type
ind...the indicator value for each species
ind_level...the Heat_requirement threshold value for calculating summed cover
iter...number of iterations
obl...allows to specify if certain abundance levels (and above) are to be reatained in every sample
rat...ratio of the number of species in the respective ecoystem type (i.e. defines sampling ratio in the bootstrap)
``` {r}
indBoot.HeatOverhang <- function(sp,abun,ind, ind_level,iter,obl,rat) {
  
  # data frame to hold the results
  heat.ind.b <- matrix(nrow=iter,ncol=length(colnames(abun)))
  colnames(heat.ind.b) <- colnames(abun)
  heat.ind.b <- as.data.frame(heat.ind.b)  
  
  # looping through the ecosystem types  
  for (j in 1:length(colnames(abun))) {
    
    # creating the species data for the respective ecosystem type
    dat <- cbind(sp,abun[,j],ind)
    dat <- dat[dat[,2]>0,]            # only species that are present in the ecoystem
    dat <- dat[!is.na(dat[,3]),]      # only species that have Heat requirement values
    dat <- as.data.frame(dat)
    
    # bootstrapping, can specify to always keep certain abundance classes (obl) and how many species to sample (rat)
    for (i in 1:iter) {
      
      speciesSample <- sample(dat$sp[dat[,2] < obl], size=round( (length(dat$sp)-length(dat$sp[dat[,2] >= (obl)])) *rat,0), replace=F)  
      dat.b <- rbind(dat[dat[,2] >= (obl),],
                     dat[match(speciesSample,dat$sp),]
      )
      dat.b[,2] <- as.numeric(dat.b[,2])
      dat.b[,3] <- as.numeric(dat.b[,3])
      colnames(dat.b)[2] <- 'abun'      
      
      # calculating the cumulative cover of species with a certain level of heat requirement and above
      if ( nrow(dat.b)>2 ) {
        
        dat.b[,2] <- dat.b[,2]/sum(dat.b[,2]) # scaling to total cover of 100%
        heat.ind.b[i,j] <- sum(dat.b[dat.b$ind>=ind_level[j],'abun'])
        
      } else {heat.ind.b[i,j] <- NA}
      
      #print(paste(i,"",j)) 
    }
    
  }
  return(heat.ind.b)
}
```

We run the bootstrap (here with 1000 iterations for now)
Here, we keep the two highest abundance classes (4/5 & 1, argument 'obl'), sampling 1/2 of the original list (argument 'rat')  
Note that I tried different values for rat (from 1/3 to 4/5)...but the results hardly changed

```{r}
heat.ind.b <- indBoot.HeatOverhang(sp=mount.dat.cov[,1],
                                   abun=mount.dat.cov[,2:28],
                                   ind=mount.dat.cov[,29],
                                   ind_level=heat.scal.tab.boot$ind_level[1:27],
                                   iter=1000,obl=4/5,rat=1/2)
```
Then we derive the limit and reference values from the bootstrapped distribution
e.g. limit value = 95% percentile, reference value = median, maximum value = 1 (i.e. the cumulative cover already reaching 1 before reaching the heat requirement threshold)

```{r error = FALSE, message = FALSE}
for (i in 1:ncol(heat.ind.b) ) {
  
  tryCatch({
    
    # lim
    heat.scal.tab.boot[i,'lim'] <- quantile(heat.ind.b[,i],probs=0.95)
    # ref
    heat.scal.tab.boot[i,'ref'] <- quantile(heat.ind.b[,i],probs=0.5)
    
  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}

```


```{r include = FALSE}
## some fixing of double NiN-type species lists

# removing the lower values (of a & b) for the NiN types with 2 species lists
heat.scal.tab.boot <- heat.scal.tab.boot[-c(7,25),]
rownames(heat.scal.tab.boot) <- 1:nrow(heat.scal.tab.boot)
heat.scal.tab.boot[c(7,24),'NiN'] <- c("T3-C1","T22-C1C3")
# cloning the species lists representative of 2 NiN-types
heat.scal.tab.boot <- rbind(heat.scal.tab.boot,
                       heat.scal.tab.boot[c(3,4,5,6,19,21,24,25),])
rownames(heat.scal.tab.boot) <- 1:nrow(heat.scal.tab.boot)
heat.scal.tab.boot[,'NiN'] <- c("T14-C1","T14-C2","T3-C3","T3-C9","T3-C2","T3-C8",
                           "T3-C1","T3-C4","T3-C7","T3-C10","T3-C13","T3-C14",
                           "T7-C1","T7-C2","T7-C3","T7-C6","T7-C8","T7-C12",
                           "T7-C13","T7-C4","T7-C7","T7-C5","T7-C10","T22-C1","T22-C2",
                           "T3-C6","T3-C12","T3-C5","T3-C11","T7-C14","T7-C9","T22-C3","T22-C4")


# add extra '-' after 'C' in heat.scal.tab.boot
heat.scal.tab.boot$NiN <- as.character(heat.scal.tab.boot$NiN)
for (i in 1:nrow(heat.scal.tab.boot) ) {
  if ( grepl('-C',heat.scal.tab.boot$NiN[i]) ) {
    xxx <- as.character(heat.scal.tab.boot$NiN[i])
    n <- which(strsplit(xxx, "")[[1]]=="C")+1
    heat.scal.tab.boot$NiN[i] <- paste(substr(xxx, 1, n-1), "-", substr(xxx, n, nchar(xxx)), sep = "")
  }
}

```

```{r}
heat.scal.tab.boot
```


## Scaling of the evaluation data to an ecological-condition index
We calculate the relative cumulative cover for species at and above the heat requirement threshold for the ANO data and evaluate these against their respective reference values (i.e. scaling according to reference value=1, limit value=0.6, and maximum value=0)
This is just technical, so I don't show the code
```{r include = FALSE}
## scaled values ####
r.s <- 1    # reference value
l.s <- 0.6  # limit value
a.s <- 0    # abscence of indicator, or indicator at maximum

## function for calculating scaled values for measured value, without truncation
scal.2 <- function() {
  # place to hold the result
  x <- numeric()
  if (maxmin < ref) {
    # values >= the reference value estimated from the linear relationship for lim < x < ref (line below)
    if (val >= ref) {x <- (l.s + (val-lim) * ( (r.s-l.s) / (ref-lim) ) )}
    # values < the reference value and >= the limit value can be deducted from the linear relationship between these two
    if (val < ref & val >= lim) {x <- (l.s + (val-lim) * ( (r.s-l.s) / (ref-lim) ) )}
    # values < the limit value and > maxmin can be deducted from the linear relationship between these two
    if (val < lim & val > maxmin) {x <- (a.s + (val-maxmin) * ( (l.s-a.s) / (lim-maxmin) ) )}
    # value equal or lower than maxmin
    if (val <= maxmin) {x <-0}
  } else {
    # values <= the reference value estimated from the linear relationship for lim < x < ref (line below)
    if (val <= ref) {x <- ( r.s - ( (r.s - l.s) * (val - ref) / (lim - ref) ) )}
    # values > the reference value and <= the limit value can be deducted from the linear relationship between these two
    if (val > ref & val <= lim) {x <- ( r.s - ( (r.s - l.s) * (val - ref) / (lim - ref) ) )}
    # values > the limit value and < maxmin can be deducted from the linear relationship between these two
    if (val > lim & val < maxmin) {x <- ( l.s - (l.s * (val - lim) / (maxmin - lim) ) )}
    # value equal og larger than maxmin
    if (val >= maxmin) {x <-0}
  }
  return(x)
  
}



#### creating dataframe to holde the results for mountains ####
ANOpkt.mount <- ANOpkt[ANOpkt$Hovedoekosystem_rute=='fjell' | ANOpkt$Hovedoekosystem_sirkel=='fjell',]
ANOpkt.mount$Hovedtype_rute <- factor(ANOpkt.mount$Hovedtype_rute)
ANOpkt.mount$Kartleggingsenhet_rute <- factor(ANOpkt.mount$Kartleggingsenhet_rute)

# choose columns for site description
colnames(ANOpkt.mount)
result.mount <- ANOpkt.mount[,c(2,3,1,4:12,15:19)]

# add columns for indicators
nvar.site <- ncol(result.mount)
result.mount[,1+nvar.site] <- NA
colnames(result.mount)[(nvar.site+1)] <- "Heat"
result.mount[,nvar.site+1] <- as.numeric(result.mount[,nvar.site+1])


## calculating scaled and non-truncated values for the indicators based on the dataset ####
for (i in 1:nrow(ANOpkt.mount) ) {  #
  tryCatch({
    print(i)
    print(paste(ANOpkt.mount$ANO_flate_ID[i]))
    print(paste(ANOpkt.mount$ANO_punkt_ID[i]))
    
    
    if (ANOpkt.mount$Hovedtype_rute[i] %in% c('T3','T7','T14','T22') ) {
      
      # Species indicator values  
      if ( length(sp[sp$ParentGlobalID==as.character(ANOpkt.mount$GlobalID[i]),'Species']) > 0 ) {
        
        
        #Heat requirement overhang
        dat <- sp[sp$ParentGlobalID==as.character(ANOpkt.mount$GlobalID[i]),c('Species','Dekning','Heat_requirement','Kartleggingsenhet_rute')]
        dat <- dat[!is.na(dat$Heat_requirement) & dat$Dekning>0,]
        
        if ( nrow(dat)>0 ) {
          
          dat$Dekning <- dat$Dekning/sum(dat$Dekning)
          val <- sum(dat[dat$Heat_requirement>=heat.scal.tab.boot[heat.scal.tab.boot$NiN==as.character(unique(dat$Kartleggingsenhet_rute)),'ind_level'],'Dekning'])
          ref <- heat.scal.tab.boot[heat.scal.tab.boot$NiN==as.character(unique(dat$Kartleggingsenhet_rute)),'ref']
          lim <- heat.scal.tab.boot[heat.scal.tab.boot$NiN==as.character(unique(dat$Kartleggingsenhet_rute)),'lim']
          maxmin <- heat.scal.tab.boot[heat.scal.tab.boot$NiN==as.character(unique(dat$Kartleggingsenhet_rute)),'maxmin']
          
          # coercing x into results.mount dataframe
          result.mount[i,'Heat'] <- scal.2() 

        } else {
          result.mount[i,'Heat'] <- 1 
        }
        
        
      }
    }
    
    
    
  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}

summary(result.mount)
# only the relevant NiN-types
result.mount <- result.mount[result.mount$Hovedtype_rute %in% c('T3','T7','T14','T22'),]
# for Heat values>1 should equal 1
result.mount[result.mount$Heat>1 & !is.na(result.mount$Heat),"Heat"] <- 1
```
Here's the result
```{r echo = FALSE}
summary(result.mount[,"Heat"])
```

And here's the result as boxplots by main ecosystem types  
T3 = Fjellhei, leeside & tundra  
T14 = Rabbe  
T7 = Snøleie  
T22 = Fjellgrashei og grastundra  
```{r echo = FALSE}
result.mount$Hovedtype_rute <- as.factor(result.mount$Hovedtype_rute)
result.mount$Hovedtype_rute <- factor(result.mount$Hovedtype_rute)

with(result.mount,plot(Hovedtype_rute,Heat))
```

Not quite unexpected it is the leesides that show the strongest signal. They are largely below the threshold value of 0.6, indicating a larger cover of species with higher Heat_requirement than expected for NiN-species-lists.















## Sjekkliste
Sjekklisten er ikke utfylt enda.

![Sjekkliste](../output/test/abc.pdf){width=100%}


